---
title: "MH3510 Assignment 1"
author: "Neo Shun Xian Nicholas"
date: "24 September 2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
   
**Goal: To explore the relationship between the amount of beta-erythoidine in an aqueous solution and the colorimeter reading of the turbidity**  
  
**X: Concentration (mg/mL)**  
**Y: Colorimeter Reading**  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Declaration of data
**Storing the data into vector X and Y for concentration and colorimeter reading respectively**

```{r storeData}
X <- c(40,50,60,70,80,90,40,60,80,50)
Y <- c(69,175,272,335,490,415,72,265,492,180)
```

## a: Fit a simple regression to the data
**Fit linear model**
```{r slr}
slr <- lm(Y~X)
# Display basic information of this model
slr
```
From the plot above, we can see that for the best fit line of the Colorimeter Reading against Concentration, the Y-intercept is -252.297 and the gradient is 8.529. i.e Y = -252.297 + 8.529*X is the best fit line.

**Plot data and the best fit line**
```{r plotLine}
# Plot data
plot(X,Y)
# Plot best fit line
abline(slr)
```
   
As shown in the above plot, it seems that the linear model of Y against X fits fairly well to the data, except with larger deviations of actual value of Y and and predicted value of Y when X equals to 80 and 90.

## b: Obtain the residuals and examine them
**Obtain and display the residual values**
```{r residuals}
# Obtain residuals
residual <- slr$res
# display the residuals value
residual
```
**Examining the residual values based on the output above**  
From the residual values, there are 3 points of predicted Y values that has significant deviations from the actual Y values with residual values -100.31, 59.978 and 61.978

## c: Comment on the adequacy of the model
**Examine the residuals by plotting**
```{r resPlot}
# Plot residual against X plot
plot(X,residual)
```
    
**Comment**   
From the above plot, we can see that the plot between the residual and Concentration(X) is **not linear**. This implies that there is **no linear relationship** between the response and predicted variable (Y_actual vs Y_predicted). Hence, the model is **not adequate** to show the relationship between the Concentration(X) and Colorimeter Reading(Y). Moreover, as shown in the plot above, it seems to imply that a **higher order polynomial term such as x^2 should be fitted** in order to show the adequacy of the model, even without the abnormal behaviour of Y when X is 80 & 90.
