---
title: "MH3510 Project Report"
author: "Neo Shun Xian Nicholas"
date: "22 October 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Background
This data is about traffic monitoring. One of the most important traffic monitoring variables is the average annual daily traffic (aadt) for a section of road or highway. It is defined as the average, over a year, of the the number of vehicles that pass through a particular section of a road each day.  
  
  
**Response Variable:**   
y: Average annual daily traffic 
  
**Predictor Variable:**  
X1: Population of country in which road section is located  
X2: Number of lanes in road section   
X3: Width of the road section (in feet)  
X4: Two-category quality variable indicating whether or not there is control of access to road section (1=access control, 2=no access control)  
**Can define X4 to be 1 if there is access control and 0 if no control**  
  
*Numerical predictor variable:* X1, X2, X3    
*Categorical predictor varable:* X4  

## Importing the dataset
```{r readData}
# read the dataset
data_raw = read.table("aadt.txt",header=FALSE)
# check data by showing the first few entries of data
head(data_raw,5)
```

## Preprocess raw data
```{r preprocess_data}
# select only the columns with the predictor and response variable
data = data.frame(y=data_raw$V1,
                  x1=data_raw$V2,
                  x2=data_raw$V3,
                  x3=data_raw$V4,
                  x4=data_raw$V5)

#defining X4 to be 1 if there is access control and 0 if there isn't
data$x4[data$x4==1] <- 0
data$x4[data$x4==2] <- 1

# take a look at the modification to the data
head(data,5)
```

## Scatter Plot Matrix
```{r scatterPlot}
# plot the scatter plot matrix
pairs(data,panel=panel.smooth)
```
  
From the plot above, there exist relations between each predictor variables (X1 to X4) and the response variable (y). 

### Model 1: Additional e^x~2~^ term in the Multi Regression Model
From the plot above, there seemed to have an exponential relation between the response variable y and predictor variable x~2~. Hence, we fit the data with an additional e^x~2~^ term: $$y=\beta_{0}+\beta_{1}X_{1}+\beta_{2}X_{2}+\beta_{3}X_{3}+\beta_{4}X_{4}+\beta_{5}e^{X_{2}}+\epsilon$$

This will be our suggested model 1.

### Fit Model 1
```{r fit1}
# fit a multiple linear regression model
mlr1 <- lm(y ~ x1+x2+x3+x4+I(exp(x2)), data=data)
summary(mlr1)
```

### Check the adequacy using t-value (standard error) and F-test between 2 models
From the result of model 1, predictor X~3~ is not significant due to very small t-value. Hence, it's coefficient may be equal to zero i.e $$H_{0}: \beta_{3} = 0$$

```{r model1}
# F-test between 2 models (full and model without predictor variable X3)
mlr1_alt <- lm(y ~ x1+x2+x4+I(exp(x2)), data=data)
# ANOVA
anova(mlr1_alt,mlr1)
```
From the above analysis, since the F-value is small, we cannot reject the null hypothesis, therefore, we remove the predictor variable X~3~ from further analysis

